Draft Final Report: Measles risk assessment, modelling, and benefit--cost analyses for New Zealand

### 2015-04-20 draftfinalreport_v3.Rnw
revisions for the final draft

New Zealand has been working towards elimination of endemic (domestic) measles virus transmission, but has suffered from small, yet significant outbreaks after measles introductions from abroad. In this draft final report we review the draft \emph {Progress Towards Measles Elimination in New Zealand - Final} report from the New Zealand Ministry of Health to the World Health Organization (WHO) Western Pacific Region. We identify additional analyses that may help understand risk of infection in New Zealand. Here we present the results of statistical analyses of risk factors for measles cases in New Zealand during outbreaks since 2007 until June 2014. We provide cost analyses for the measles outbreaks in New Zealand, and include modelling of measles outbreaks, including pre- and post-vaccination scenarios, based on the numbers of na\"{\i}ve people at the District Health Board (DHB) and national level. We provide benefit--cost analyses using the results from those model simulations, along with a number of alternative vaccination strategies to achieve different vaccination coverage levels. Our key findings are:

\begin{itemize}
\item The \emph {Progress Towards Measles Elimination in New Zealand - Final} report was of high quality and contained substantial information and useful analyses.
\item The regression analyses suggest that age is a particularly strong risk factor for measles.  Increased wealth (NZDep 1--3) was a risk factor and people of European ethnicity make up the majority of cases. However, our analyses also highlight other groups that are at greater risk of measles on a per capita basis. Pacific people are at greater risk per capita. However, this was driven by young Pacific children, as 2--24 year old Pacific people are less at risk than European and Maori, as are 5--17 year old Asian children.
\item Of the 11\% of the overall population of New Zealand that are currently considered to be immunologically na\"{\i}ve, it was estimated that approximately 28\% would require additional vaccination to ensure measles does not persist, leading to an additional 131,500 vaccinations nationally.
\item New Zealand is at risk of frequent measles importation due to travel and endemic measles elsewhere in the globe. Peak overall overseas travel rates, and thus presumably risk from measles importation, is in December. However, New Zealander and immigrant or non--New Zealander travel is otherwise out of phase, with peak travel for New Zealanders during winter, and non--New Zealanders during summer.
\item Analyses of outbreak data suggest that measles basic reproduction number ($R_v$, the number of secondary infections in a partially vaccinated population) values often include or exceed one, including 2013--2014 outbreaks. As analysed from data until June, the last outbreak to end in 2014 had an $R_v$ well above 1 (mean estimate 3.66, range 3.4--3.92). The range of estimates for $R_v$ between 2009--2014 was 0.18--3.92. These analyses suggest improved vaccination is a requisite to prevent measles persisting in the population and that large outbreaks could occur.
\item The cost of 187 measles cases in 2014 is estimated to be approximately \$1,041,186 due to earnings lost relating to cases and contacts, case management and hospitalisation costs.
\item The mean wage loss per measles case is estimated to be approximately \$839.
\item The mean cost per measles case attending hospital is estimated to be \$1,877 per case, and approximately 17\% of measles cases attend hospital
\item The mean public health service cost per case is \$1,765 excluding hospitalisation costs.
\item The average number of contacts requiring quarantine was 2.11, requiring 7.3 days of quarantine on average, at a cost of \$170.4 per day.
\item The benefit--cost (B/C) ratio analyses suggest additional vaccination is extremely beneficial financially (B/C $>$1), with vaccine costs required to exceed approximately \$8,214 per vaccine before the costs exceed the benefits.
\item Following catch up vaccination to ensure $R_v$ is < 1 measles introductions to New Zealand with a median size of 2 cases were predicted by simulation. Thus, typically individual cases and secondary cases will be the expectation. However, the mean outbreak size of 61 cases was predicted, because outbreak sizes of tens, hundreds, and very rarely up to many thousands of cases are predicted among the 1000 simulations following importation, despite $R_v$ being one and the outbreak predicted to die out. Thus, increased vaccination beyond the 28\% of the currently 11\% na\'{\i}ve population required may be useful to prevent these rare but costly events.
\end{itemize}

@@

@@

\vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
  print(xtb,floating = FALSE)
@
\caption{Summary of the regression model results using \autoref{eq:reg} and a quasipoisson error structure, for measles risk factors for cases from 2007-June 2014}
\label{table:regression_riskfactors}
\end{table}

@@

The majority of measles cases are among people of European ancestry (\autoref{table:percap} and \autoref{fig:cases_all}). Indeed, the majority of cases were in more wealthy European 5--17 year olds (\autoref{fig:cases_all}). However, per capita rates for this large age class were lower than in the very young (0--1 year olds, \autoref{fig:percap_all}). People of Pacific origin are over-represented as measles cases on a per capita basis ($\beta$ = 1.22, standard error (SE) = 0.21, p--value < 0.0001; \autoref{table:regression_riskfactors}), and in particular the 0--1 year old less wealthy (NZDep 8--10) (\autoref{fig:percap_all}). Note however that generally those people in NZDep levels 4--10 are less represented on a per capita basis (NZDep 4--7 $\beta$ = -0.26, SE = 0.09, p--value 0.009; NZDep 8--10 $\beta$ = -0.3, SE = 0.1, p--value == 0.006; \autoref{table:regression_riskfactors}), and in European and Maori ethnicities there is a trend opposite to that in than Pacific people, with the more wealthy more likely to be cases (\autoref{fig:percap_all} and \autoref{fig:cases_all}).

@@

\begin{figure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{vacc_age.pdf}
\end{center}
\caption{Age of vaccination of vaccinated measles cases, 2007- June 2014. Recommended vaccination age in months is shown for the measles, mumps, rubella (MMR) dose 1 (12 months) and MMR dose 2 (60 months)}
\label{fig:vaccstat}
\end{figure}

\begin{figure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{dob_vacc_unvac.pdf}
\end{center}
\caption{Year of birth of unvaccinated of measles cases, 2007- June 2014}
\label{fig:yearandvac}
\end{figure}

\begin{figure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{dob_vacc_dose.pdf}
\end{center}
\caption{Year of birth of unvaccinated of measles cases, 2007- June 2014. The year of birth for those vaccinated with measles, mumps, and rubella (MMR) dose 1 is shown in the upper panel, MMR dose 2 the lower panel.}
\label{fig:yearandvacf}
\end{figure}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12006)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12006[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2006 birth cohort vaccinated against MMR1 across the 2013 area units. See \autoref{fig:fig12006}.}
\label{table:tab12006}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22006)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22006[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2006 birth cohort vaccinated against MMR2 across the 2013 area units. See \autoref{fig:fig22006}.}
\label{table:tab22006}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12007)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12007[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2007 birth cohort vaccinated against MMR1 across the 2013 area units. See \autoref{fig:fig12007}.}
\label{table:tab12007}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22007)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22007[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2007 birth cohort vaccinated against MMR2 across the 2013 area units. See \autoref{fig:fig22007}.}
\label{table:tab22007}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12008)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12008[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2008 birth cohort vaccinated against MMR1 across the 2013 area units. See \autoref{fig:fig12008}.}
\label{table:tab12008}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22008)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22008[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2008 birth cohort vaccinated against MMR2 across the 2013 area units. See \autoref{fig:fig22008}.}
\label{table:tab22008}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12009)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12009[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2009 birth cohort vaccinated against MMR1 across the 2013 area units. See \autoref{fig:fig12009}.}
\label{table:tab12009}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22009)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22009[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2009 birth cohort vaccinated against MMR2 across the 2013 area units. See \autoref{fig:fig22009}.}
\label{table:tab22009}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12010)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12010[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2010 birth cohort vaccinated against MMR1 across the 2013 area units. See \autoref{fig:fig12010}.}
\label{table:tab12010}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22010)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22010[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2010 birth cohort vaccinated against MMR2 across the 2013 area units. See \autoref{fig:fig22010}.}
\label{table:tab22010}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12011)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12011[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2011 birth cohort vaccinated against MMR1 across the 2013 area units. See \autoref{fig:fig12011}.}
\label{table:tab12011}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22011)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22011[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2011 birth cohort vaccinated against MMR2 across the 2013 area units. See \autoref{fig:fig22011}.}
\label{table:tab22011}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12012)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12012[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2012 birth cohort vaccinated against MMR1 across the 2013 area units. See \autoref{fig:fig12012}.}
\label{table:tab12012}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22012)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22012[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2012 birth cohort vaccinated against MMR2 across the 2013 area units. See \autoref{fig:fig22012}.}
\label{table:tab22012}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12013)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12013[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2006 birth cohort vaccinated against MMR1 across the 2013 area units. See \autoref{fig:fig12013}.}
\label{table:tab12013}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22013)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22013[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2006 birth cohort vaccinated against MMR2 across the 2013 area units. See \autoref{fig:fig22013}.}
\label{table:tab22013}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12014)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12014[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2014 birth cohort vaccinated against MMR1 across the 2013 area units. See \autoref{fig:fig12014}.}
\label{table:tab12014}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22014)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22014[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2014 birth cohort vaccinated against MMR2 across the 2013 area units. See \autoref{fig:fig22014}.}
\label{table:tab22014}
\end{table}

@@

Plotting the lower level data available from the NIR database highlights the issue of surveillance data requiring appropriate denominator data, as these analyses show that there is a lack of correspondence between the national census data and the NIR data (\autoref{fig:naive_au}, \autoref{fig:naive_au_prop})

>>>>

New Zealand has been working towards elimination of endemic (domestic) measles virus transmission, but has suffered from small, yet significant outbreaks after measles introductions from abroad. In this draft final report we review the draft \emph {Progress Towards Measles Elimination in New Zealand - Final} report from the New Zealand Ministry of Health to the World Health Organization (WHO) Western Pacific Region. We identify additional analyses that may help understand risk of infection in New Zealand. Here we present the results of statistical analyses of risk factors for measles cases in New Zealand during outbreaks since 2007 until 12 June 2014. We provide cost analyses for the measles outbreaks in New Zealand, and include modelling of measles outbreaks, including pre- and post-vaccination scenarios, based on the numbers of na\"{\i}ve people at the District Health Board (DHB) and national level. We provide benefit--cost analyses using the results from those model simulations, along with a number of alternative vaccination strategies to achieve different vaccination coverage levels. Our key findings are:

\begin{itemize}
\item The \emph {Progress Towards Measles Elimination in New Zealand - Final} report was of high quality and contained substantial information and useful analyses.
\item The regression analyses suggest that age is a particularly strong risk factor for measles.  Increased wealth (NZDep 1--3) was a risk factor and people of European ethnicity make up the majority of cases. However, our analyses also highlight other groups that are at greater risk of measles on a per capita basis. Pacific people are at greater risk per capita. However, this was driven by young Pacific children, as 2--24 year old Pacific people are less at risk than European and Maori, as are 5--17 year old Asian children.
\item Of the 11\% of the overall population of New Zealand that are currently considered to be immunologically na\"{\i}ve, it was estimated that approximately 28\% would require additional vaccination to ensure measles does not persist, leading to an additional 128,034 vaccinations nationally.
\item New Zealand is at risk of frequent measles importation due to travel and endemic measles elsewhere in the globe. Peak overall overseas travel rates, and thus presumably risk from measles importation, is in December. However, New Zealander and immigrant or non--New Zealander travel is otherwise out of phase, with peak travel for New Zealanders during winter, and non--New Zealanders during summer.
\item Analyses of outbreak data suggest that measles basic reproduction number ($R_v$, the number of secondary infections in a partially vaccinated population) values often include or exceed one, including 2013--2014 outbreaks. As analysed from data until 12 June 2014, the last outbreak to end in 2014 had an $R_v$ well above 1 (mean estimate 3.66, range 3.4--3.92). The range of estimates for $R_v$ between 2009--2014 was 0.18--3.92. These analyses suggest improved vaccination is a requisite to prevent measles persisting in the population and that large outbreaks could occur.
\item The cost of the first 187 measles cases in 2014 is estimated to be approximately \$1,041,186 due to earnings lost relating to cases and contacts, case management and hospitalisation costs.
\item The mean wage loss per measles case is estimated to be approximately \$839.
\item The mean cost per measles case attending hospital is estimated to be \$1,877 per case, and approximately 17\% of measles cases attend hospital
\item The mean public health service cost per case is \$1,765 excluding hospitalisation costs.
\item The average number of contacts requiring quarantine was 2.11, requiring 7.3 days of quarantine on average, at a cost of \$170.4 per day.
\item The benefit--cost (B/C) ratio analyses suggest additional vaccination is extremely beneficial financially (B/C $>$1), with vaccine costs required to exceed approximately \$5,365 per vaccine before the costs exceed the benefits for all DHBs.
\item Following catch up vaccination to ensure $R_v$ is < 1 measles introductions to New Zealand with a median size of 2 cases were predicted by simulation. Thus, typically individual cases and secondary cases will be the expectation. However, the mean outbreak size of 61 cases was predicted, because outbreak sizes of tens, hundreds, and very rarely up to many thousands of cases are predicted among the 1000 simulations following importation, despite $R_v$ being one and the outbreak predicted to die out. Thus, increased vaccination beyond the 28\% of the currently 11\% na\'{\i}ve population required may be useful to prevent these rare but costly events.
\end{itemize}

@@

R0 = 12.8;
x0max = 0.3;
Nx0 = 200;
P = (1:Nx0)*x0max/(Nx0+1);
x0 = P/(1-exp(-R0*P));
Rv = R0*x0;
Pop = c(436350,206000,482180,283700,469300,151700,138380,98196,162560,137000,151690,
        55620,297420,43650,109750,359310,41112,525550,32151,60120,4241739);
Naive = c(51673,20511,50995,32406,55085,15471,15078,10470,17196,12960,14793,5199,
          31389,4727,11378,39087,3901,57904,3172,6025,459420);
RvDHB = R0*Naive/Pop;

library(akima)
PDHB = aspline(x=Rv,y=P,xout=RvDHB);

# The number of cases per DHB
FSDHB = PDHB$y*Naive;
# The number of vaccinations short
VDHB = Naive-Pop/R0;

Z<-data.frame(nrows=21,ncols=5)
for (i in 1:21){
Z[i,1] = Pop[i];
Z[i,2] = Naive[i];
Z[i,3] = round(FSDHB[i]);
Z[i,4] = round(VDHB[i]);
Z[i,5] = round(round(VDHB[i])/Naive[i],2)}
colnames(Z)<-c("Population","Naive","Outbreak","Vaccination","PC")

write.csv(Z,"dhb_vacc.csv",row.names=F)

@@

\vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
  print(xtb,floating = FALSE)
@
\caption{Summary of the regression model results using \autoref{eq:reg} and a quasipoisson error structure, for measles risk factors for cases from 2007-12 June 2014}
\label{table:regression_riskfactors}
\end{table}

@@

The majority of measles cases are among people of European ancestry (\autoref{table:percap} and \autoref{fig:cases_all}). Indeed, the majority of cases were in more wealthy European 5--17 year olds (\autoref{fig:cases_all}). However, per capita rates for this large age class were lower than in the very young (0--1 year olds, \autoref{fig:percap_all}). People of Pacific origin are over-represented as measles cases on a per capita basis ($\beta$ = 1.22, standard error (SE) = 0.21, p--value < 0.0001; \autoref{table:regression_riskfactors}), and in particular the 0--1 year old less wealthy (NZDep 8--10) (\autoref{fig:percap_all}). Note however that generally those people in NZDep levels 4--10 are less represented on a per capita basis (NZDep 4--7 $\beta$ = -0.26, SE = 0.09, p--value 0.009; NZDep 8--10 $\beta$ = -0.3, SE = 0.1, p--value = 0.006; \autoref{table:regression_riskfactors}), and in European and Maori ethnicities there is a trend opposite to that in than Pacific people, with the more wealthy more likely to be cases (\autoref{fig:percap_all} and \autoref{fig:cases_all}).

@@

\begin{figure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{vacc_age.pdf}
\end{center}
\caption{Age of vaccination of vaccinated measles cases, 2007- 12 June 2014. Recommended vaccination age in months is shown for the measles, mumps, rubella (MMR) dose 1 (12 months) and MMR dose 2 (60 months)}
\label{fig:vaccstat}
\end{figure}

\begin{figure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{dob_vacc_unvac.pdf}
\end{center}
\caption{Year of birth of unvaccinated of measles cases, 2007- 12 June 2014}
\label{fig:yearandvac}
\end{figure}

\begin{figure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{dob_vacc_dose.pdf}
\end{center}
\caption{Year of birth of unvaccinated of measles cases, 2007- 12 June 2014. The year of birth for those vaccinated with measles, mumps, and rubella (MMR) dose 1 is shown in the upper panel, MMR dose 2 the lower panel.}
\label{fig:yearandvacf}
\end{figure}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12006)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12006[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2006 birth cohort vaccinated against MMR1 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig12006}.}
\label{table:tab12006}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22006)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22006[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2006 birth cohort vaccinated against MMR2 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig22006}.}
\label{table:tab22006}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12007)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12007[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2007 birth cohort vaccinated against MMR1 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig12007}.}
\label{table:tab12007}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22007)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22007[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2007 birth cohort vaccinated against MMR2 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig22007}.}
\label{table:tab22007}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12008)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12008[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2008 birth cohort vaccinated against MMR1 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig12008}.}
\label{table:tab12008}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22008)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22008[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2008 birth cohort vaccinated against MMR2 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig22008}.}
\label{table:tab22008}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12009)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12009[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2009 birth cohort vaccinated against MMR1 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig12009}.}
\label{table:tab12009}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22009)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22009[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2009 birth cohort vaccinated against MMR2 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig22009}.}
\label{table:tab22009}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12010)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12010[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2010 birth cohort vaccinated against MMR1 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig12010}.}
\label{table:tab12010}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22010)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22010[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2010 birth cohort vaccinated against MMR2 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig22010}.}
\label{table:tab22010}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12011)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12011[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2011 birth cohort vaccinated against MMR1 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig12011}.}
\label{table:tab12011}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22011)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22011[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2011 birth cohort vaccinated against MMR2 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig22011}.}
\label{table:tab22011}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12012)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12012[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2012 birth cohort vaccinated against MMR1 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig12012}.}
\label{table:tab12012}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22012)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22012[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2012 birth cohort vaccinated against MMR2 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig22012}.}
\label{table:tab22012}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12013)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12013[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2006 birth cohort vaccinated against MMR1 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig12013}.}
\label{table:tab12013}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22013)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22013[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2006 birth cohort vaccinated against MMR2 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig22013}.}
\label{table:tab22013}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t12014)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t12014[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2014 birth cohort vaccinated against MMR1 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig12014}.}
\label{table:tab12014}
\end{table}

@@

 \vspace{5mm} %5mm vertical space
\begin{table}
<<include=FALSE,echo=FALSE,results=tex>>=
colnames(t22014)<-c("Min","1st Q","Median","Mean","3rd Q","Max","NA")
latex(t22014[1:7], file="", table.env=FALSE,rowname=NULL)
@
\caption{Summary of the proportion of 2014 birth cohort vaccinated against MMR2 across the 2013 area units. Area units are aggregations of meshblock data. See \autoref{fig:fig22014}.}
\label{table:tab22014}
\end{table}

@@

Plotting the lower level data from the 2013 area unit data, available from the NIR database, shows that there is heterogeneity at the local level and an obvious temporal lag, as the proportion vaccinated declines with birth cohort (\autoref{fig:mmr_nir}), though recent vaccination levels are consistently high for MMR1. For full details, see \autoref{fig:fig12006} (MMR1, 2006), \autoref{fig:fig22006} (MMR2, 2006), \autoref{fig:fig12007} (MMR1, 2007), \autoref{fig:fig22007} (MMR2, 2007), \autoref{fig:fig12008} (MMR1, 2008), \autoref{fig:fig22008} (MMR2, 2008), \autoref{fig:fig12009} (MMR1, 2009), \autoref{fig:fig22009} (MMR2, 2009), \autoref{fig:fig12010} (MMR1, 2010), \autoref{fig:fig22010} (MMR2, 2010), \autoref{fig:fig12011} (MMR1, 2011), \autoref{fig:fig22011} (MMR2, 2011), \autoref{fig:fig12012} (MMR1, 2012), \autoref{fig:fig22012} (MMR2, 2012), \autoref{fig:fig12013} (MMR1, 2013), \autoref{fig:fig22013} (MMR2, 2013), \autoref{fig:fig12014} (MMR1, 2014), \autoref{fig:fig22014} (MMR2, 2014), \autoref{table:tab12006} (MMR1, 2006), \autoref{table:tab22006} (MMR2, 2006), \autoref{table:tab12007} (MMR1, 2007), \autoref{table:tab22007} (MMR2, 2007), \autoref{table:tab12008} (MMR1, 2008), \autoref{table:tab22008} (MMR2, 2008), \autoref{table:tab12009} (MMR1, 2009), \autoref{table:tab22009} (MMR2, 2009), \autoref{table:tab12010} (MMR1, 2010), \autoref{table:tab22010} (MMR2, 2010), \autoref{table:tab12011} (MMR1, 2011), \autoref{table:tab22011} (MMR2, 2011), \autoref{table:tab12012} (MMR1, 2012), \autoref{table:tab22012} (MMR2, 2012), \autoref{table:tab12013} (MMR1, 2013), \autoref{table:tab22013} (MMR2, 2013), \autoref{table:tab12014} (MMR1, 2014), \autoref{table:tab22014} (MMR2, 2014) and the supplementary files.

<<echo=FALSE,include=FALSE,results=hide>>=
MMR1<-c(0.9604,0.9621,0.971,0.9722,0.9667,0.9643,0.9667,0.6333,0)
MMR2<-c(0.9167,0.9099,0.9048,0.8889,0.7059,0.0345,0.0041,0,0)
res_bar<-cbind(MMR1,MMR2)
rownames(res_bar)<-c("2006","2007","2008","2009","2010","2011","2012","2013","2014")
@

\begin{figure}
\begin{center}
<<fig=TRUE,echo=FALSE>>=
barplot(t(res_bar),beside=T,ylim=c(0,1),xlab="Birth year",ylab="Proportion vaccinated",legend=colnames(res_bar),
        args.legend = list(x = "topright",bty="n",inset=c(0.1,0.1)))
abline(h=0.90,col="red",lty=3,lwd=4)
abline(h=0.95,col="orange",lty=3,lwd=4)
@
\end{center}
\caption{Proportions vaccinated by birth cohort using the NIR database. 95\% and 90\% thresholds are shown by orange and red dotted lines respectively}
\label{fig:mmr_nir}
\end{figure}

The correlation between the percentage vaccinated against MMR1 and MMR2 for the 2004 2013 area units is shown in figure \autoref{fig:cor_nir}.
